---
title: "1. Formating"
output: html_document
date: "2024-10-24"
fig_width: 6 
fig_height: 10 
---

This is the first step of the data analysis pipeline. 

In this code we run basic statistics about our sample and we add new variables to the data file.

1. We begin by loading the required packages.

```{r packages, echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(dplyr)
library(patchwork)
library(ggplot2)
library(openxlsx)
library(lme4)
library(lmerTest)
library(stats)
library(emmeans)
library(tidyboot)
library(faintr)
library(brms)
library(stringr)
library(dabestr)
library(ez)
library(MBESS)
```

2. We compute basic statistics about the sample for the method section.

```{r Sample, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE}
data_participants <- "C:/Users/grandjeamari/Documents/Travail/UCLouvain/PhD/Projet/Projet-Saccades/Miscellaneous/Infos_Participants_ET.xlsx"
data_participants <- read_excel(data_participants)

# Age of the subjects
data_participants %>%
  summarise(mean_Age = mean(Age, na.rm = TRUE),
            sd_Age = sd(Age, na.rm = TRUE))

# Number of women
data_participants %>%
  summarise(Women_Count = sum(Genre == "Femme", na.rm = TRUE))

# Handedness of the participants
data_participants %>%
  summarise(`Right-Handed_Count` = sum(str_detect(Latéralité, "^D"), na.rm = TRUE))

# frACT 
data_participants %>%
  filter(!is.na(`frACT (Log)`)) %>%
  separate(`frACT (Log)`, into = c("value_logmar", "dummy"), sep = ", ") %>%
  mutate(
    value_logmar = as.numeric(gsub("[()]", "", value_logmar)),
  ) %>%
  summarise(
    mean_logmar = mean(value_logmar, na.rm = TRUE),
    sd_logmar = sd(value_logmar, na.rm = TRUE))

# Benton
data_participants %>% 
  mutate(Benton = as.numeric(Benton)) %>% 
  summarise(mean_benton = mean(Benton, na.rm = TRUE),
            sd_benton = sd(Benton, na.rm = TRUE))

```


3. Now, we fetch the data and we create new variables that will allow us to assess whether the saccades were accurate.
The datafile is an excel sheet obtained after assembling all the edf files together via Matlab.

```{r Formating, echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}

#### Specify the path to your Excel file 
excel_file5 <- "C:/Users/grandjeamari/Documents/Travail/UCLouvain/PhD/Projet/Projet-Saccades/Analysis/Data/Eyelink/Data_50.xlsx"

# Read the Excel data into a data frame
excel_data <- read_excel(excel_file5)

# Create a new variable to write the position of the target
excel_data <- excel_data %>%
  mutate(correct_location = case_when(
    grepl("hori", tolower(Subjects)) ~ ifelse(Target_LocationX < 0, "left", "right"),
    grepl("vert", tolower(Subjects)) ~ ifelse(Target_LocationY < 0, "down", "up"),
    TRUE ~ NA_character_
  ))

# Create a new variable to check if the saccade was accurate
excel_data <- excel_data %>%
  mutate(Accuracy = case_when(
    grepl("hori", tolower(Subjects)) & correct_location == HorizSaccDir1 ~ 1,
    grepl("vert", tolower(Subjects)) & correct_location == VertSaccDir1 ~ 1,
    TRUE ~ 0
  ))

#### Create a new file with the added variables 

# Specify the file path for the new Excel file
output_excel_file <- "C:/Users/grandjeamari/Documents/Travail/UCLouvain/PhD/Projet/Projet-Saccades/Analysis/Data/Eyelink/Output_50.xlsx"

# Write the dataframe to an Excel file
write.xlsx(excel_data, file = output_excel_file, sheetName = "Sheet1", row.names = FALSE)

```

--> Once all of this is done, we are good to go to step number 2.