---
title: "2. Descriptive Statistics"
output: html_document
date: "2024-10-24"
fig_width: 6 
fig_height: 10 
---
This is the second step of the data analysis pipeline. 

In this code we run descriptive statistics about the accuracy and latency of the saccades.

1. We begin by loading the required packages .
```{r packages, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(dplyr)
library(patchwork)
library(ggplot2)
library(openxlsx)
library(lme4)
library(lmerTest)
library(stats)
library(emmeans)
library(tidyboot)
library(faintr)
library(brms)
library(stringr)
library(dabestr)
library(ez)
library(MBESS)
```

2. We load the data, remove the training trials and create a new dataframe.

```{r Removing training, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE}

### Open the file and remove training 
excel_data <- read_excel("C:/Users/grandjeamari/Documents/Travail/UCLouvain/PhD/Projet/Projet-Saccades/Analysis/Data/Eyelink/Output_50.xlsx")

excel_data$TRIALID <- as.numeric(excel_data$TRIALID)

excel_data_no_training <- excel_data %>% 
  filter(TRIALID > 19)

# now we retake the same name for simplicity purpose
excel_data <- excel_data_no_training

```

3. We assess whether all subjects have a reasonable amount of valid trials.
NB: This is just a sanity check, we can decide to exclude or not subjects based on the quantity of invalid trials.

```{r Checking valid saccades , echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE}

# Check by subject 
valid_counts <- excel_data %>%
  group_by(Subjects) %>%
  summarize(subject_num_valid = mean(Valid == 1, na.rm = TRUE) * 100)
print(valid_counts)

invalid_subjects <- valid_counts %>% 
  filter(subject_num_valid <= 75)
print(invalid_subjects$Subjects)

# Check by target
valid_counts <- excel_data %>%
  group_by(Target) %>%
  summarize(target_num_valid = mean(Valid == 1, na.rm = TRUE) * 100)
print(valid_counts)
# Removing participants that have below 75% of valid trials

#excel_data <- excel_data %>% 
#    filter(!str_detect(Subjects, "^BED|^BEL|^FRL|^MAL"))

#### Check how many trials where removed ####
cat("The retrieval of invalid saccades led to the deletion of",
    round(mean(excel_data$Valid == 0, na.rm = TRUE) * 100, 2),
    "% of the trials\n")

```

4. We create a clean dataframe to facilitate further analysis.

```{r creating new dataframes, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE}

# Create clean_df 
clean_data <- excel_data %>% 
  dplyr::select(Subjects, Target, Accuracy, Valid, Latency1, Distractor, correct_location, TRIALID) %>%
  mutate(Subjects = substr(Subjects, 1, 8)) %>%
  mutate(Accuracy=as.character(Accuracy)) %>%
  filter(Valid == "1") %>% 
  mutate(condition=case_when(Accuracy == "1" & Target == "faces" ~ "face",
                             Accuracy == "1" & Distractor == "faces" ~ "vehi",
                             Accuracy == "0" & Target == "vehicule" ~ "face",
                             Accuracy == "0" & Distractor == "vehicule" ~ "vehi")) %>%
  separate(Subjects, into=c("subject","meridian")) %>% 
  mutate(Target=case_when(Target == "faces" ~ "Faces",
                          Target == "vehicule" ~ "Vehicles"),
         meridian=case_when(meridian == "hori" ~ "Horizontal",
                       meridian == "vert" ~ "Vertical")) %>% 
  mutate(Accuracy = as.numeric(Accuracy))

save(clean_data, file = "clean_data.rda")

# Creating a df with only accurate saccades
clean_data_correct <- clean_data %>% 
  filter(Accuracy == "1")

```

5. We look at the saccadic accuracy. We can then decide whether we want to exclude participants. 

```{r checking accuracy, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE}
#### Check how many participant we have ####
nsuj <- length(unique(substr(clean_data$subject, 1, 3)))
print(paste("number of subjects:" , nsuj))

# Check if participant have poor accuracy ####
mean_acc <- clean_data %>%
  dplyr::group_by(subject, Target, meridian) %>%
  dplyr::summarize(mean_acc = mean(Accuracy, na.rm = TRUE),
            sd_acc = sd(Accuracy, na.rm = TRUE)) %>%
  dplyr::mutate(message = ifelse(mean_acc < 0.6, paste("Mean accuracy below 0.6 for", subject), NA))

print(mean_acc, n = Inf, na.print = "")

# Remove participants that have at least one condition below 60% accuracy
#clean_data <- clean_data %>%
#  filter(!str_detect(subject, "^FRL|^MAL|^PIB|^MRB|^STC"))

#excel_data <- excel_data %>%
#  filter(!str_detect(Subjects, "^FRL|^MAL|^PIB|^MRB|^STC"))

```

6. We start the analysis with saccadic accuracy. 
As we will do a bayesian analysis on saccadic reaction time (SRT), we only go with a frequentist analysis for accuracies.

```{r Accuracy analysis, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE}

##### Descriptive analysis #####
# Perform ANOVA analysis
results <- ezANOVA(data = clean_data,
                   dv = .(Accuracy),
                   wid = .(subject),
                   within = .(Target, meridian),
                   type = 3,
                   detailed = TRUE,
                   return_aov = TRUE)

# Calculate partial eta squared
results$ANOVA$partialetasquared <- results$ANOVA$SSn/(results$ANOVA$SSn+results$ANOVA$SSd)

# Print results
print(results$ANOVA)


# Global accuracy for faces
clean_data %>%
  filter(Target == "Faces") %>%
  summarize(mean_accuracy_faces = mean(Accuracy, na.rm = TRUE),
            sd_accuracy_faces = sd(Accuracy, na.rm = TRUE))

# Global accuracy for vehicles
clean_data %>%
  filter(Target == "Vehicles") %>%
  summarize(mean_accuracy_vehicles = mean(Accuracy, na.rm = TRUE),
            sd_accuracy_vehicles = sd(Accuracy, na.rm = TRUE))

# Global horizontal accuracy
clean_data %>%
  filter(grepl("Horizontal", meridian)) %>%
  summarize(mean_accuracy_hori = mean(Accuracy, na.rm = TRUE),
            sd_accuracy_hori = sd(Accuracy, na.rm = TRUE))

# Global vertical accuracy
clean_data %>%
  filter(grepl("Vertical", meridian)) %>%
  summarize(mean_accuracy_vert = mean(Accuracy, na.rm = TRUE),
            sd_accuracy_vert = sd(Accuracy, na.rm = TRUE))

```

7. Here we describe the mean accuracy and latency of the saccades for the horiontal meridian.

```{r horizontal analysis, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE}

## Accuracy ##

hori_accuracy_faces <- clean_data %>%
  filter(Target == "Faces" & grepl("Horizontal", meridian)) %>%
  summarize(mean_accuracy_faces_hori = mean(Accuracy, na.rm = TRUE),
            sd_accuracy_faces_hori = sd(Accuracy, na.rm = TRUE))

hori_accuracy_vehicule <- clean_data %>%
  filter(Target == "Vehicles" & grepl("Horizontal", meridian)) %>%
  summarize(mean_accuracy_vehicles_hori = mean(Accuracy, na.rm = TRUE),
            sd_accuracy_vehicles_hori = sd(Accuracy, na.rm = TRUE))

hori_accuracy_general <- clean_data %>% 
  filter(grepl("Horizontal", meridian)) %>%
  summarize(mean_accuracy_hori = mean(Accuracy, na.rm = TRUE),
            sd_accuracy_hori = sd(Accuracy, na.rm = TRUE))

# Print the results
print(hori_accuracy_general)
print(hori_accuracy_faces)
print(hori_accuracy_vehicule)

# left side only
left_accuracy_general <- clean_data %>% 
  filter(grepl("left", correct_location)) %>%
  summarize(mean_accuracy_left = mean(Accuracy, na.rm = TRUE),
            sd_accuracy_left = sd(Accuracy, na.rm = TRUE))

left_accuracy_faces <- clean_data %>%
  filter(Target == "Faces" & grepl("left", correct_location)) %>%
  summarize(mean_accuracy_faces_left = mean(Accuracy, na.rm = TRUE),
            sd_accuracy_faces_left = sd(Accuracy, na.rm = TRUE))

left_accuracy_vehicule <- clean_data %>%
  filter(Target == "Vehicles" & grepl("left", correct_location)) %>%
  summarize(mean_accuracy_vehicles_left = mean(Accuracy, na.rm = TRUE),
            sd_accuracy_vehicles_left = sd(Accuracy, na.rm = TRUE))

# right side only
right_accuracy_general <- clean_data %>% 
  filter(grepl("right", correct_location)) %>%
  summarize(mean_accuracy_right = mean(Accuracy, na.rm = TRUE),
            sd_accuracy_right = sd(Accuracy, na.rm = TRUE))

right_accuracy_faces <- clean_data %>%
  filter(Target == "Faces" & grepl("right", correct_location)) %>%
  summarize(mean_accuracy_faces_right = mean(Accuracy, na.rm = TRUE),
            sd_accuracy_faces_right = sd(Accuracy, na.rm = TRUE))

right_accuracy_vehicule <- clean_data %>%
  filter(Target == "Vehicles" & grepl("right", correct_location)) %>%
  summarize(mean_accuracy_vehicles_right = mean(Accuracy, na.rm = TRUE),
            sd_accuracy_vehicles_right = sd(Accuracy, na.rm = TRUE))

print(left_accuracy_general)
print(left_accuracy_faces)
print(left_accuracy_vehicule)
print(right_accuracy_general)
print(right_accuracy_faces)
print(right_accuracy_vehicule)

## Latency ##

hori_latency_faces <- clean_data_correct %>%
  filter(Target == "Faces" & grepl("Horizontal", meridian)) %>%
  summarize(mean_latency_faces_hori = mean(Latency1, na.rm = TRUE),
            sd_latency_faces_hori = sd(Latency1, na.rm = TRUE))

hori_latency_vehicule <- clean_data_correct %>%
  filter(Target == "Vehicles" & grepl("Horizontal", meridian)) %>%
  summarize(mean_latency_vehicles_hori = mean(Latency1, na.rm = TRUE),
            sd_latency_vehicles_hori = sd(Latency1, na.rm = TRUE))

hori_latency_general <- clean_data_correct %>%
  filter(grepl("Horizontal", meridian)) %>%
  summarize(mean_latency_hori = mean(Latency1, na.rm = TRUE),
            sd_latency_hori = sd(Latency1, na.rm = TRUE))

# Print the results
print(hori_latency_general)
print(hori_latency_faces)
print(hori_latency_vehicule)

# left side only
left_latency_general <- clean_data_correct %>% 
  filter(grepl("left", correct_location)) %>%
  summarize(mean_latency_left = mean(Latency1, na.rm = TRUE),
            sd_latency_left = sd(Latency1, na.rm = TRUE))

left_latency_faces <- clean_data_correct %>%
  filter(Target == "Faces" & grepl("left", correct_location)) %>%
  summarize(mean_latency_left = mean(Latency1, na.rm = TRUE),
            sd_latency_left = sd(Latency1, na.rm = TRUE))

left_latency_vehicule <- clean_data_correct %>%
  filter(Target == "Vehicles" & grepl("left", correct_location)) %>%
  summarize(mean_latency_left = mean(Latency1, na.rm = TRUE),
            sd_latency_left = sd(Latency1, na.rm = TRUE))

#right side only
right_latency_general <- clean_data_correct %>% 
  filter(grepl("right", correct_location)) %>%
  summarize(mean_latency_right = mean(Latency1, na.rm = TRUE),
            sd_latency_right = sd(Latency1, na.rm = TRUE))

right_latency_faces <- clean_data_correct %>%
  filter(Target == "Faces" & grepl("right", correct_location)) %>%
  summarize(mean_latency_faces_right = mean(Latency1, na.rm = TRUE),
            sd_latency_faces_right = sd(Latency1, na.rm = TRUE))

right_latency_vehicule <- clean_data_correct %>%
  filter(Target == "Vehicles" & grepl("right", correct_location)) %>%
  summarize(mean_latency_vehicles_right = mean(Latency1, na.rm = TRUE),
            sd_latency_vehicles_right = sd(Latency1, na.rm = TRUE))

print(left_latency_general)
print(left_latency_faces)
print(left_latency_vehicule)
print(right_latency_general)
print(right_latency_faces)
print(right_latency_vehicule)

```

8. Here we describe the mean accuracy and latency of the saccades for the vertical meridian.

```{r vertical analysis, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE}

## Accuracy ##

vert_accuracy_faces <- clean_data %>%
  filter(Target == "Faces" & grepl("Vertical", meridian)) %>%
  summarize(mean_accuracy_faces_vert = mean(Accuracy, na.rm = TRUE),
            sd_accuracy_faces_vert = sd(Accuracy, na.rm = TRUE))

vert_accuracy_vehicule <- clean_data %>%
  filter(Target == "Vehicles" & grepl("Vertical", meridian)) %>%
  summarize(mean_accuracy_vehicles_vert = mean(Accuracy, na.rm = TRUE),
            sd_accuracy_vehicles_vert = sd(Accuracy, na.rm = TRUE))

vert_accuracy_general <- clean_data %>% 
  filter(grepl("Vertical", meridian)) %>%
  summarize(mean_accuracy_vert = mean(Accuracy, na.rm = TRUE),
            sd_accuracy_vert = sd(Accuracy, na.rm = TRUE))

# Print the results
print(vert_accuracy_general)
print(vert_accuracy_faces)
print(vert_accuracy_vehicule)

# up side only
up_accuracy_general <- clean_data %>% 
  filter(grepl("up", correct_location)) %>%
  summarize(mean_accuracy_up = mean(Accuracy, na.rm = TRUE),
            sd_accuracy_up = sd(Accuracy, na.rm = TRUE))

up_accuracy_faces <- clean_data %>%
  filter(Target == "Faces" & grepl("up", correct_location)) %>%
  summarize(mean_accuracy_faces_up = mean(Accuracy, na.rm = TRUE),
            sd_accuracy_faces_up = sd(Accuracy, na.rm = TRUE))

up_accuracy_vehicule <- clean_data %>%
  filter(Target == "Vehicles" & grepl("up", correct_location)) %>%
  summarize(mean_accuracy_vehicles_up = mean(Accuracy, na.rm = TRUE),
            sd_accuracy_vehicule_up = sd(Accuracy, na.rm = TRUE))

# down side only
down_accuracy_general <- clean_data %>% 
  filter(grepl("down", correct_location)) %>%
  summarize(mean_accuracy_down = mean(Accuracy, na.rm = TRUE),
            sd_accuracy_down = sd(Accuracy, na.rm = TRUE))

down_accuracy_faces <- clean_data %>%
  filter(Target == "Faces" & grepl("down", correct_location)) %>%
  summarize(mean_accuracy_faces_down = mean(Accuracy, na.rm = TRUE),
            sd_accuracy_faces_down = sd(Accuracy, na.rm = TRUE))

down_accuracy_vehicule <- clean_data %>%
  filter(Target == "Vehicles" & grepl("down", correct_location)) %>%
  summarize(mean_accuracy_vehicles_down = mean(Accuracy, na.rm = TRUE),
            sd_accuracy_vehicles_down = sd(Accuracy, na.rm = TRUE))

print(up_accuracy_general)
print(up_accuracy_faces)
print(up_accuracy_vehicule)
print(down_accuracy_general)
print(down_accuracy_faces)
print(down_accuracy_vehicule)

## Latency ##

vert_latency_faces <- clean_data_correct %>%
  filter(Target == "Faces" & grepl("Vertical", meridian)) %>%
  summarize(mean_latency_faces_vert = mean(Latency1, na.rm = TRUE),
            sd_latency_faces_vert = sd(Latency1, na.rm = TRUE))

vert_latency_vehicule <- clean_data_correct %>%
  filter(Target == "Vehicles" & grepl("Vertical", meridian)) %>%
  summarize(mean_latency_vehicles_vert = mean(Latency1, na.rm = TRUE),
            sd_latency_vehicles_vert = sd(Latency1, na.rm = TRUE))

vert_latency_general <- clean_data_correct %>%
  filter(grepl("Vertical", meridian)) %>%
  summarize(mean_latency_vert = mean(Latency1, na.rm = TRUE),
            sd_latency_vert = sd(Latency1, na.rm = TRUE))

# Print the results
print(vert_latency_general)
print(vert_latency_faces)
print(vert_latency_vehicule)

# up side only
up_latency_general <- clean_data_correct %>% 
  filter(grepl("up", correct_location)) %>%
  summarize(mean_latency_up = mean(Latency1, na.rm = TRUE),
            sd_latency_u = sd(Latency1, na.rm = TRUE))

up_latency_faces <- clean_data_correct %>%
  filter(Target == "Faces" & grepl("up", correct_location)) %>%
  summarize(mean_latency_faces_up = mean(Latency1, na.rm = TRUE),
            sd_latency_faces_up = sd(Latency1, na.rm = TRUE))

up_latency_vehicule <- clean_data_correct %>%
  filter(Target == "Vehicles" & grepl("up", correct_location)) %>%
  summarize(mean_latency_vehicles_up = mean(Latency1, na.rm = TRUE),
            sd_latency_vehicles_up = sd(Latency1, na.rm = TRUE))

# down side only
down_latency_general <- clean_data_correct %>% 
  filter(grepl("down", correct_location)) %>%
  summarize(mean_latency_down = mean(Latency1, na.rm = TRUE),
            sd_latency_down = sd(Latency1, na.rm = TRUE))

down_latency_faces <- clean_data_correct %>%
  filter(Target == "Faces" & grepl("down", correct_location)) %>%
  summarize(mean_latency_faces_down = mean(Latency1, na.rm = TRUE),
            sd_latency_faces_down = sd(Latency1, na.rm = TRUE))

down_latency_vehicule <- clean_data_correct %>%
  filter(Target == "Vehicles" & grepl("down", correct_location)) %>%
  summarize(mean_latency_vehicles_down = mean(Latency1, na.rm = TRUE),
            sd_latency_vehicles_down = sd(Latency1, na.rm = TRUE))

print(up_latency_general)
print(up_latency_faces)
print(up_latency_vehicule)
print(down_latency_general)
print(down_latency_faces)
print(down_latency_vehicule)


```

9. Finaly, we create some plots to visualize the data.

```{r data visualization, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE}
# Boxplot
clean_data_2 <- clean_data %>% 
  dplyr::select(subject, meridian, Accuracy, Latency1, condition) %>%
  filter(Accuracy == "1") %>%
  group_by(subject, meridian, condition) %>%
  dplyr::summarise(mean_lat=mean(Latency1)) %>%
  mutate(pos = case_when(condition=="face" ~ 1.4,
                          condition=="vehi" ~ 1.6),
         jitter = jitter(pos,factor=1)) %>% 
  dplyr::rename(Target = condition)


plt <- clean_data_2 %>%
  ggplot(aes(x = pos, y = mean_lat, color=Target, fill = Target)) +
  geom_boxplot(aes(colour = Target, fill=Target), 
               outlier.shape = NA, alpha = 0.5, width = 0.1) +
  geom_point(aes(x = jitter, y = mean_lat, fill = Target, group=subject),
             size = 6, shape = 20) +
  geom_line(aes(x=jitter,group=subject), color="black", alpha=.1)+
  stat_boxplot(aes(group=Target),geom="errorbar", width=.1)+
  facet_wrap(~meridian) +
  #geom_flat_violin(aes(fill = condition),
                   #position = position_nudge(x = .13, y = 0), 
                   #adjust = 1, trim = FALSE, alpha = .75)+
  scale_fill_manual(values = c("orange", "blue"),
                    labels = c("Face", "Vehicle")) +
  scale_colour_manual(values = c("orange", "blue"),
                      labels = c("Face", "Vehicle")) +
  labs(y = "Saccadic reaction time (ms)",
       x = 'Meridian',
       title = "Distribution of saccadic reaction time") +
  stat_summary(
    aes(group = Target),
    fun = "mean",
    geom = "point",
    size = 6, # Adjust size as needed
    color = "black",
    shape = 15,
    alpha = 0.75,
    show.legend = FALSE
  ) +
  stat_summary(
    aes(group = meridian),
    fun = "mean",
    geom = "line",
    color = "black",
    alpha = 0.75, 
    show.legend = FALSE
  )+
  theme_classic() +
  theme(axis.title.y = element_text(size = 25),
        axis.text.x = element_blank(),
        axis.text.y = element_text(size = 20),
        axis.title.x = element_blank(),
        axis.line.x=element_blank(),
        strip.background = element_blank(),
        strip.text = element_text(size = 25),
        plot.title = element_text(hjust = 0.5, size = 30))+
  scale_x_discrete(expand = c(0.3, 0))+
  ylim(120,380)

print(plt)

ggsave("plt.png", plt, dpi=600, width = 9, height = 7.5)

# Let's also plot the accuracy
clean_data_3 <- clean_data %>% 
  dplyr::select(subject, meridian, Accuracy, Latency1, condition) %>%
  group_by(subject, meridian, condition) %>%
  dplyr::summarise(mean_acc=mean(Accuracy)) %>%
  mutate(pos = case_when(condition=="face" ~ 1.4,
                         condition=="vehi" ~ 1.6),
         jitter = jitter(pos,factor=1)) %>% 
  dplyr::rename(Target = condition)

plt2 <- clean_data_3 %>%
  ggplot(aes(x = pos, y = mean_acc, color=Target, fill = Target)) +
  geom_boxplot(aes(colour = Target, fill=Target), 
               outlier.shape = NA, alpha = 0.5, width = 0.1) +
  geom_point(aes(x = jitter, y = mean_acc, fill = Target, group=subject),
             size = 6, shape = 20) +
  geom_line(aes(x=jitter,group=subject), color="black", alpha=.1)+
  stat_boxplot(aes(group=Target),geom="errorbar", width=.1)+
  facet_wrap(~meridian) +
  #geom_flat_violin(aes(fill = condition),
  #position = position_nudge(x = .13, y = 0), 
  #adjust = 1, trim = FALSE, alpha = .75)+
  scale_fill_manual(values = c("orange", "blue"),
                    labels = c("Face", "Vehicle")) +
  scale_colour_manual(values = c("orange", "blue"),
                      labels = c("Face", "Vehicle")) +
  labs(y = "Saccadic reaction time (ms)",
       x = 'Meridian',
       title = "Distribution of saccadic accuracy") +
  stat_summary(
    aes(group = Target),
    fun = "mean",
    geom = "point",
    size = 6, # Adjust size as needed
    color = "black",
    shape = 15,
    alpha = 0.75,
    show.legend = FALSE
  ) +
  stat_summary(
    aes(group = meridian),
    fun = "mean",
    geom = "line",
    color = "black",
    alpha = 0.75, 
    show.legend = FALSE
  )+
  theme_classic() +
  theme(axis.title.y = element_text(size = 25),
        axis.text.x = element_blank(),
        axis.text.y = element_text(size = 20),
        axis.title.x = element_blank(),
        axis.line.x=element_blank(),
        strip.background = element_blank(),
        strip.text = element_text(size = 25),
        plot.title = element_text(hjust = 0.5, size = 30))+
  scale_x_discrete(expand = c(0.3, 0))+
  ylim(0,1)

print(plt2)

ggsave("plt2.png", plt, dpi=600, width = 9, height = 7.5)

```
In this section, we create new dataframe that we will use for later correlations

```{r data visualization, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE}

raw_face_adv_hori <-clean_data_2 %>% 
  filter(meridian %in% "Horizontal") %>% 
  group_by(subject) %>%
  summarize(hori_adv = sum(mean_lat[Target == "vehi"]) - sum(mean_lat[Target == "face"])) %>%
  dplyr::select(subject, hori_adv)


raw_face_adv_vert <-clean_data_2 %>% 
  filter(meridian %in% "Vertical") %>% 
  group_by(subject) %>%
  summarize(vert_adv = sum(mean_lat[Target== "vehi"]) - sum(mean_lat[Target == "face"])) %>%
  dplyr::select(subject, vert_adv)

raw_face_adv <- data.frame(subject = raw_face_adv_hori$subject, hori_adv = raw_face_adv_hori$hori_adv, vert_adv = raw_face_adv_vert$vert_adv)

save(raw_face_adv, file = "raw_face_adv.rda")

# We create this df to later check the correlations using these scores
# We start by looking at all the saccades for the specific target regardless of meridian
raw_face <- clean_data_2 %>% 
  group_by(subject) %>%
  summarize(meanSRT = mean(mean_lat[Target == "face"])) %>%
  dplyr::select(subject, meanSRT)

raw_vehi <- clean_data_2 %>% 
  group_by(subject) %>%
  summarize(meanSRT = mean(mean_lat[Target == "vehi"])) %>%
  dplyr::select(subject, meanSRT)

save(raw_face, file = "raw_face.rda")
save(raw_vehi, file = "raw_vehi.rda")

# Now we also split by meridian
raw_face_hori <-clean_data_2 %>% 
  filter(meridian %in% "Horizontal") %>% 
  group_by(subject) %>%
  summarize(meanSRT = sum(mean_lat[Target == "face"])) %>%
  dplyr::select(subject, meanSRT)

raw_face_vert <-clean_data_2 %>% 
  filter(meridian %in% "Vertical") %>% 
  group_by(subject) %>%
  summarize(meanSRT = sum(mean_lat[Target == "face"])) %>%
  dplyr::select(subject, meanSRT)

raw_vehi_hori <-clean_data_2 %>% 
  filter(meridian %in% "Horizontal") %>% 
  group_by(subject) %>%
  summarize(meanSRT = sum(mean_lat[Target == "vehi"])) %>%
  dplyr::select(subject, meanSRT)

raw_vehi_vert <-clean_data_2 %>% 
  filter(meridian %in% "Vertical") %>% 
  group_by(subject) %>%
  summarize(meanSRT = sum(mean_lat[Target == "vehi"])) %>%
  dplyr::select(subject, meanSRT)

save(raw_face_hori, file = "raw_face_hori.rda")
save(raw_face_vert, file = "raw_face_vert.rda")
save(raw_vehi_hori, file = "raw_vehi_hori.rda")
save(raw_vehi_vert, file = "raw_vehi_vert.rda")

# Count how many subjects obtained a bigger score for hori_adv
num_subjects_bigger_hori_adv <- raw_face_adv %>%
  filter(hori_adv >= 0 & vert_adv >= 0) %>% 
  summarise(num_subjects = sum(hori_adv > vert_adv))

# Print the number of subjects
print(num_subjects_bigger_hori_adv)

# Set the color of the points based on specific values (orange if adv for faces, otherwise blue if adv for vehicles)
raw_face_adv$colour_condition <- ifelse(raw_face_adv$hori_adv > raw_face_adv$vert_adv, "#E46F34", "#EE9A09")
raw_face_adv$colour_condition[!(raw_face_adv$hori_adv > raw_face_adv$vert_adv & raw_face_adv$hori_adv > 0 & raw_face_adv$vert_adv > 0)] <- "grey"


raw_scatterplt <- raw_face_adv %>%
  ggplot() + 
  ggtitle("Face advantage (Vehicle - Face)") +
  labs(x = "Horizontal meridian", y = "Vertical meridian") +
  geom_dotplot(aes(x = hori_adv, y = vert_adv, group = subject), 
               binaxis='y', stackdir='center', 
               colour = raw_face_adv$colour_condition, 
               fill = raw_face_adv$colour_condition, 
               dotsize = 2, binwidth = .75)+ 
  geom_abline(intercept = 0, slope = 1)+
  geom_abline(hjust = 0, vjust = 0, intercept = 0, slope = 0, linetype = "dashed", color = "#EE9A09", alpha = 1) +  # Add a horizontal line
  geom_vline(xintercept = 0, linetype = "dashed", color = "#E46F34", alpha = 1) +  # Vertical line
  guides(linetype = FALSE) +
  scale_y_continuous(limits = c(-40, 50), breaks = seq(-40, 50, by = 5)) +
  scale_x_continuous(limits = c(-40, 50), breaks = seq(-40, 50, by = 5)) +
  theme(panel.grid = element_blank(),
        aspect.ratio=1, #axis.text = element_text(size = 30),
        panel.border = element_rect(colour = "black", fill=NA, size=3),
        panel.background = element_rect(fill = "white", colour = "black"))
print(raw_scatterplt)

ggsave("myplot3.png", raw_scatterplt, dpi=600, width = 9, height = 6.5)

```

--> Once all of this is done, we are good to go to step number 3.
